Informing policy via dynamic models: Eliminating cholera in Haiti

### 2022-11-30 ms.Rnw
Change model3 model to fit Iinit and hurricane parameters

%% Template for the submission to:
%%   The Annals of Applied Statistics [AOAS]

@@

@@

# Create vectors for the unit and shared parameters
unit_specific_names <- c("betaB", "foi_add")
shared_param_names <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k", "sigma", "rho", "gamma"
)

mod3_n_params <- length(unit_specific_names) * 10 + length(shared_param_names)

@@

set_rw_1 <- function(x) {
  if (grepl('^foi_add[[:digit:]]+$', x) || grepl('^betaB[[:digit:]]+$', x)) {
    return(0.01 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.025)))
  } else {
    return(0.01)
  }
}

set_rw_2 <- function(x) {
  if (grepl('^foi_add[[:digit:]]+$', x) || grepl('^betaB[[:digit:]]+$', x)) {
    return(0.003 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.0075)))
  } else {
    return(0.003)
  }
}

set_rw_3 <- function(x) {
  if (grepl('^foi_add[[:digit:]]+$', x) || grepl('^betaB[[:digit:]]+$', x)) {
    return(0.001 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.002)))
  } else {
    return(0.001)
  }
}

@@

# Search hyper-parameters for each run-level and search
if (RUN_LEVEL == 1){
  SEARCH1 = list(
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 100,
    NREPS_EVAL = 5,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.05,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 5,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 100,
    SPAT_REGRESSION = 0.05,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 2L
} else if (RUN_LEVEL == 2) {
  SEARCH1 = list(
    NBPF = 25,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 36,
    NP_EVAL = 1000,
    NREPS_EVAL = 12,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  SEARCH2 = list(
    TOP_N = 6,
    NBPF = 25,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 12,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 25,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 12,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  N_SEARCHES <- 1L
} else if (RUN_LEVEL == 3) {

  SEARCH1 = list(
    NBPF = 100,
    NP = 5000,
    SPAT_REGRESSION = 0.05,
    NREPS = 144,
    NP_EVAL = 5000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE
  )

  SEARCH2 = list(
    TOP_N = 12,
    NBPF = 100,
    NP = 5000,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 5000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE
  )

  SEARCH3 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 5000,
    SPAT_REGRESSION = 0.05,
    NREPS = 8,
    NP_EVAL = 5000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIK_MAT = TRUE
  )

  N_SEARCHES <- 3L
} else {
  stop("Run level must be 1, 2 or 3.")
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/haiti3_coupled_fit.rds")) {
  haiti3_fit <- readRDS("model3/run_level_3/haiti3_coupled_fit.rds")
} else {
  haiti3_fit <- bake(
    file = paste0("model3/", rl_dir, "haiti3_coupled_fit.rds"), {
      fit_coupled_haiti3(
        nsearches = N_SEARCHES,
        search1 = SEARCH1,
        search2 = SEARCH2,
        search3 = SEARCH3,
        ncores = cores,
        search_rho = TRUE,
        search_gamma = TRUE,
        search_Binit = FALSE
      )
    }
  )
}

@@

for (p in shared_param_names) {
  pat <- paste0("^", p, "[[:digit:]]+$")
  p3s[p] <-  mean(p3[grepl(pat, names(p3))])
}

>>>>

%% Based on the template for the submission to:
%%   The Annals of Applied Statistics [AOAS]

@@

if (RUN_LEVEL == 3) {
  # Print session info into a txt file. Only update if RUN_LEVEL == 3, which
  # is when the final results are calculated.
  cat(
    utils::capture.output(sessionInfo()),
    file = "sessionInfo.txt",
    sep="\n"
  )
}

@@

# Create vectors for the unit and shared parameters
unit_specific_names <- c("betaB", "foi_add", "aHur", "hHur", "Iinit")
shared_param_names <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k"
)

# betaB + foi_add for all 10 units +
# aHur3, aHur9, hHur3, hHur9 +
# Iinit3, Iinit7, Iinit9 +
# shared_parameters
mod3_n_params <- 2 * 10 + 7 + length(shared_param_names)

@@

set_rw_1 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.01)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.015, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.25)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.015)
  } else {
    return(0)
  }
}

set_rw_2 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.003)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.005, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.125)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.005)
  } else {
    return(0)
  }
}

set_rw_3 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.0025)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.005, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.125)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.003)
  } else {
    return(0)
  }
}

@@

# Search hyper-parameters for each run-level and search
if (RUN_LEVEL == 1){
  SEARCH1 = list(
    NBPF = 15,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 36,
    NP_EVAL = 500,
    NREPS_EVAL = 3,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 9,
    NBPF = 10,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 4,
    NP_EVAL = 500,
    NREPS_EVAL = 3,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 10,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 500,
    NREPS_EVAL = 3,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 3L
} else if (RUN_LEVEL == 2) {
  SEARCH1 = list(
    NBPF = 100,
    NP = 1000,
    SPAT_REGRESSION = 0.05,
    NREPS = 72,
    NP_EVAL = 1000,
    NREPS_EVAL = 8,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  SEARCH2 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 1000,
    SPAT_REGRESSION = 0.05,
    NREPS = 4,
    NP_EVAL = 1000,
    NREPS_EVAL = 9,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 100,
    NP = 1000,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 9,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  N_SEARCHES <- 3L
} else if (RUN_LEVEL == 3) {

  SEARCH1 = list(
    NBPF = 100,
    NP = 2000,
    SPAT_REGRESSION = 0.05,
    NREPS = 144,
    NP_EVAL = 2000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIKE_MAT = FALSE
  )

  SEARCH2 = list(
    TOP_N = 12,
    NBPF = 100,
    NP = 2000,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 2000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIK_MAT = FALSE
  )

  SEARCH3 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 2000,
    SPAT_REGRESSION = 0.05,
    NREPS = 8,
    NP_EVAL = 2000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIK_MAT = TRUE
  )

  N_SEARCHES <- 3L
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/haiti3_coupled_fit.rds")) {
  haiti3_fit <- readRDS("model3/run_level_3/haiti3_coupled_fit.rds")
} else {
  haiti3_fit <- bake(
    file = paste0("model3/", rl_dir, "haiti3_coupled_fit.rds"), {
      fit_coupled_haiti3(
        nsearches = N_SEARCHES,
        search1 = SEARCH1,
        search2 = SEARCH2,
        search3 = SEARCH3,
        search_rho = FALSE,
        search_gamma = FALSE,
        search_Iinit = TRUE,
        search_hur = TRUE,
        ncores = cores
      )
    }
  )
}

@@

for (p in shared_param_names) {
  p3s[p] <- p3[paste0(p, 1)]  # They are the same for 1:10
}
